### 2020-07-04 chapters/abstract.tex
Small corrections

%Approach/Solution and %Main contributions
In this thesis, we design a method that makes use of lightweight streaming aggregations and distribution divergence functions to alert about deviations in data patterns in real-time, while in the presence of large volumes of high velocity, highly skewed and seasonal data. We develop a two-phased and two-windowed method: first we perform a batch analysis on a reference window and then a stream analysis over a sliding streaming window. We aggregate the contents of both windows using an approximate histogram aggregation based on Exponential Moving Averages. We do this for each event and for each of its fields, also denominated features in our context. We measure the divergence between both reference and sliding window histograms for each feature with the Jensen–Shannon Divergence function, but other distance functions could be used. For each feature, we find out the probability value of each divergence measure and apply the Holm-Bonferroni multiple test correction to find out which probability values are lowest and their associated features. We generate alerts based on a user-defined probability threshold, for each feature.

We evaluate our method through a series of tests, some with synthetic datasets and some with real data. We further split the tests into single and multi-feature analysis. Our method accurately detected the introduced anomalies in the synthetic dataset experiments while maintaining high throughput, for single and multi feature analysis. However, experiments with real data were not as accurate. Despite not knowing the specific reasons for that, we defer further investigation to future work and formulate a set of hypothesis that might explain it and hence are worthy of pursuing next. Our set of experiments supports the claim that sliding window aggregations and distribution divergence methods can be combined to detect data pattern shifts in streaming scenarios with constant time and memory complexity regarding window size.

>>>>

%Approach/Solution and %Main contributions
In this thesis, we design a method that makes use of lightweight streaming aggregations and distribution divergence functions to alert about deviations in data patterns in real-time, while in the presence of large volumes of high velocity, highly skewed and seasonal data. We develop a two-phased and two-windowed method: first, we perform batch analysis on a reference window and then a stream analysis over a sliding streaming window. We aggregate the contents of both windows using an approximate histogram aggregation based on Exponential Moving Averages. We do this for each event and each of its fields also denominated features in our context. We measure the divergence between both reference and sliding window histograms for each feature with the Jensen–Shannon Divergence function, but other distance functions could be used. For each feature, we find out the probability value of each divergence measure and apply the Holm-Bonferroni multiple test correction to find out which probability values are lowest and their associated features. We generate alerts based on a user-defined probability threshold, for each feature.

We evaluate our method through a series of tests, some with synthetic datasets and some with real data. We further split the tests into single and multi-feature analysis. Our method accurately detected the introduced anomalies in the synthetic dataset experiments while maintaining high throughput, for single and multi-feature analysis. However, experiments with real data were not as accurate. Despite not knowing the specific reasons for that, we defer further investigation to future work and formulate a set of hypotheses that might explain it and hence are worthy of pursuing next. Our set of experiments supports the claim that sliding window aggregations and distribution divergence methods can be combined to detect data pattern shifts in streaming scenarios with constant time and memory complexity regarding window size.

### 2020-07-04 chapters/background.tex
Small corrections

In this chapter we discuss fundamental concepts that will help us identify the set of relevant techniques that are suitable to address the problem to be formulated and tackled in this thesis. In Section \ref{sec:stream-superset} we introduce batch and stream processing as two large scale data processing techniques. In Section \ref{sec:windows} we review windowing mechanisms that essentially turn unbounded data streams to finite datasets. In Section \ref{sec:aggregations} we introduce aggregations as a whole. Then, we specifically discuss sliding window aggregations, algorithms used for its computation and aggregation properties. In Section \ref{sec:emas} we present Exponential Moving Averages as aggregators that replace the sliding window contents for a lighter representation, providing an approximate estimation of the aggregation value.

@@

Stream processing, on the other hand, refers to the processing of data streams. A data stream is essentially a continuous or unbounded dataset. This data needs to be processed sequentially and incrementally on a event-by-event basis. Streaming data varies widely, from logs generated by customers using mobile or web applications, sensor data from Internet-of-Things (IoT) devices to credit card transactions. These examples of infinite datasets contrast with batch processing which works on finite ones. This distinction is important since some aggregations are very hard and even impossible to compute with 100\% accuracy for several use cases in an efficient manner on unbounded datasets. An example of such is counting distinct elements that flow through the data stream. An immediate solution would be to create a regular set of items or strings and have its length be the number of distinct elements processed. However, this approach is not very efficient and has a linear memory usage growth regarding the number of distinct elements. Therefore, as the number of distinct elements approaches large orders of magnitude we run out of memory.

In order to provide a streaming solution, Apache Spark created the Spark Streaming \cite{SparkStreaming} module. Being a batch system in its core, Spark implements streaming as micro-batching \cite{SparkStreamingPaper}. This approach handles streaming data by grouping incoming events into very small batches. Then, in order to process the entire defined window of data, it joins all micro-batches into a larger one and performs the desired computations. However, this creates an artificial barrier that does not truly exist in the streaming context. More than conceptually different from a true stream processing system, processing a stream in a batch fashion, even if in small ones, presents some issues. First, since we are indeed working with batches, results will never be updated in real-time per event being processed. For example, if we use a batch size of 100 events, the system would only produce accurate results every 100 events. To handle system inactivity usually these micro-batch systems also have the notion of updating the aggregations with incomplete batches, if no event has entered the system after a pre-specified time period has elapsed. As such, micro-batching was created as a natural implementation of stream processing using the available batch processing systems of the time, but can not be considered true stream processing.

The Venn diagram in Figure \ref{fig:stream-superset} \cite{batch-is-a-special-case-of-streaming} illustrates the relation of both processing techniques. A data stream is unbounded or infinite while a bounded data stream has a beginning and an end. Using these definitions, the concept of an unbounded data stream contains that of a bounded dataset. That is, from an infinite stream, it is possible to fix a left and right limits in order to obtain a bounded and finite dataset. This relation is visible in Figure \ref{fig:unbounded-bounded-superset} from the same blog post.

@@

% boundless aggregations no need for windows
Computations over an unbounded dataset are possible. For example, computing the \textit{average} of an event's field for all incoming events is computationally trivial: we need only to store the total number of events and the current total sum of all events arriving to the system. However, is this boundless average over the stream of events useful? While in some scenarios it may be, more often than not we would like to introduce some boundaries into our computations. For instance, consider the monitoring of machines spread out across multiple racks in a large data center and its temperatures. Assume that temperatures measured across all racks are sent to a monitoring system in the form of a data stream. The average temperature of a machine over an infinite period (\textit{i.e.} since the machine was deployed) is of little value for detecting if the equipment is currently overheating, and to trigger the appropriate action to prevent a failure. In this scenario, an average of the temperatures per rack and for the last 5 minutes allows for more actionable context, in this case identifying a likely-to-fail rack. To obtain such insight requires computing the average over the last 5 minutes' worth of data, effectively defining a begin and end --- \textit{i.e.} all the data between the current timestamp in minutes \textit{t} and \textit{t-5}. By creating such a conceptual range, we realize that in practice we need a bounded dataset. Partitioning an infinite data stream to obtain such a finite dataset is made through the use of windows.

@@

Windows can also be characterized by how they move across data, \textit{i.e.}, how the step $S$ is defined. The three most common ways are: true sliding, stepping or tumbling windows. A true sliding window is a window with an unitary step, i.e., where $S = 1$. At each step, the window advances one unit, which may be one event for tuple-based steps, or one time-unit whether that is in milliseconds, seconds, days, weeks or another magnitude of time, measured based on the system clock or event time. Such a true sliding movement can be observed in Figure \ref{fig:sliding-window} with a tuple-based window of size \textit{W} of three tuples and unitary step.

@@

On the previous Section, we presented Subtract-On-Evict as an improvement over Recalculate-From-Scratch based on the idea of aggregation invertibility. The question then becomes if there are other aggregation properties besides invertibility that must be taken into account when analyzing general stream processing algorithms.

@@

We talked about sliding window aggregations and algorithms to compute them. We observed that the contents of the sliding window must be stored in memory in order to evict the oldest event and discount its effect from the aggregation when a new one is inserted. In this Section we analyze Exponential Moving Averages (EMAs). EMAs process each event only upon insertion. Since there is no eviction operation to perform, we do not need to store the whole window of data.

Exponential Moving Averages (also known as Exponential Weighted Moving Averages) \cite{EMA-Everett2011, EMA-Hunter, EMA-MarcusB} are a type of reduce functions applied to a sequence of elements, one at a time in a orderly fashion, to compute a weighted average, using different weights for each element. These weights decrease exponentially with the age of the observations. In other words, EMAs give higher weights to more recent data points. Figures \ref{fig:window-weights} and \ref{fig:ema-weights} show the weights applied to past events. In a typical sliding window, events either have weight of one (inside the window) or zero (outside the window) whereas EMAs give exponentially smaller weights to older events.

@@

Definition \ref{def:ema} introduces the recurrence formula for time-based EMAs whereas Definition \ref{def:tuple-ema} works for tuple-based ones. In this Thesis we chose to work with tuple-based EMAs due to its simplicity, since it does not require keeping track of timestamp variations, but our approach can be adjusted to use time-based ones.

@@

>>>>

In this chapter, we discuss fundamental concepts that will help us identify the set of relevant techniques that are suitable to address the problem to be formulated and tackled in this thesis. In Section \ref{sec:stream-superset} we introduce batch and stream processing as two large scale data processing techniques. In Section \ref{sec:windows} we review windowing mechanisms that essentially turn unbounded data streams to finite datasets. In Section \ref{sec:aggregations} we introduce aggregations as a whole. Then, we specifically discuss sliding window aggregations, algorithms used for its computation and aggregation properties. In Section \ref{sec:emas} we present Exponential Moving Averages as aggregators that replace the sliding window contents for a lighter representation, providing an approximate estimation of the aggregation value.

@@

Stream processing, on the other hand, refers to the processing of data streams. A data stream is essentially a continuous or unbounded dataset. This data needs to be processed sequentially and incrementally on an event-by-event basis. Streaming data varies widely, from logs generated by customers using mobile or web applications, sensor data from Internet-of-Things (IoT) devices to credit card transactions. These examples of infinite datasets contrast with batch processing which works on finite ones. This distinction is important since some aggregations are very hard and even impossible to compute with 100\% accuracy for several use cases in an efficient manner on unbounded datasets. An example of such is counting distinct elements that flow through the data stream. An immediate solution would be to create a regular set of items or strings and have its length be the number of distinct elements processed. However, this approach is not very efficient and has a linear memory usage growth regarding the number of distinct elements. Therefore, as the number of distinct elements approaches large orders of magnitude we run out of memory.

In order to provide a streaming solution, Apache Spark created the Spark Streaming \cite{SparkStreaming} module. Being a batch system in its core, Spark implements streaming as micro-batching \cite{SparkStreamingPaper}. This approach handles streaming data by grouping incoming events into very small batches. Then, to process the entire defined window of data, it joins all micro-batches into a larger one and performs the desired computations. However, this creates an artificial barrier that does not truly exist in the streaming context. More than conceptually different from a true stream processing system, processing a stream in a batch fashion, even if in small ones, presents some issues. First, since we are indeed working with batches, results will never be updated in real-time per event being processed. For example, if we use a batch size of 100 events, the system would only produce accurate results every 100 events. To handle system inactivity usually these micro-batch systems also have the notion of updating the aggregations with incomplete batches, if no event has entered the system after a pre-specified time period has elapsed. As such, micro-batching was created as a natural implementation of stream processing using the available batch processing systems of the time, but can not be considered true stream processing.

The Venn diagram in Figure \ref{fig:stream-superset} \cite{batch-is-a-special-case-of-streaming} illustrates the relation of both processing techniques. A data stream is unbounded or infinite while a bounded data stream has a beginning and an end. Using these definitions, the concept of an unbounded data stream contains that of a bounded dataset. That is, from an infinite stream, it is possible to fix a left and right limits to obtain a bounded and finite dataset. This relation is visible in Figure \ref{fig:unbounded-bounded-superset} from the same blog post.

@@

% boundless aggregations no need for windows
Computations over an unbounded dataset are possible. For example, computing the \textit{average} of an event's field for all incoming events is computationally trivial: we need only to store the total number of events and the current total sum of all events arriving in the system. However, is this boundless average over the stream of events useful? While in some scenarios it may be, more often than not we would like to introduce some boundaries into our computations. For instance, consider the monitoring of machines spread out across multiple racks in a large data center and its temperatures. Assume that temperatures measured across all racks are sent to a monitoring system in the form of a data stream. The average temperature of a machine over an infinite period (\textit{i.e.} since the machine was deployed) is of little value for detecting if the equipment is currently overheating, and to trigger the appropriate action to prevent a failure. In this scenario, an average of the temperatures per rack and for the last 5 minutes allows for a more actionable context, in this case identifying a likely-to-fail rack. To obtain such insight requires computing the average over the last 5 minutes' worth of data, effectively defining a begin and end --- \textit{i.e.} all the data between the current timestamp in minutes \textit{t} and \textit{t-5}. By creating such a conceptual range, we realize that in practice we need a bounded dataset. Partitioning an infinite data stream to obtain such a finite dataset is made through the use of windows.

@@

Windows can also be characterized by how they move across data, \textit{i.e.}, how the step $S$ is defined. The three most common ways are: true sliding, stepping or tumbling windows. A true sliding window is a window with a unitary step, i.e., where $S = 1$. At each step, the window advances one unit, which may be one event for tuple-based steps, or one time-unit whether that is in milliseconds, seconds, days, weeks or another magnitude of time, measured based on the system clock or event time. Such a true sliding movement can be observed in Figure \ref{fig:sliding-window} with a tuple-based window of size \textit{W} of three tuples and unitary step.

@@

In the previous section, we presented Subtract-On-Evict as an improvement over Recalculate-From-Scratch based on the idea of aggregation invertibility. The question then becomes if there are other aggregation properties besides invertibility that must be taken into account when analyzing general stream processing algorithms.

@@

We talked about sliding window aggregations and algorithms to compute them. We observed that the contents of the sliding window must be stored in memory to evict the oldest event and discount its effect from the aggregation when a new one is inserted. In this section, we analyze Exponential Moving Averages (EMAs). EMAs process each event only upon insertion. Since there is no eviction operation to perform, we do not need to store the whole window of data.

Exponential Moving Averages (also known as Exponential Weighted Moving Averages) \cite{EMA-Everett2011, EMA-Hunter, EMA-MarcusB} are a type of reduce functions applied to a sequence of elements, one at a time in an orderly fashion, to compute a weighted average, using different weights for each element. These weights decrease exponentially with the age of the observations. In other words, EMAs give higher weights to more recent data points. Figures \ref{fig:window-weights} and \ref{fig:ema-weights} show the weights applied to past events. In a typical sliding window, events either have a weight of one (inside the window) or zero (outside the window) whereas EMAs give exponentially smaller weights to older events.

@@

Definition \ref{def:ema} introduces the recurrence formula for time-based EMAs whereas Definition \ref{def:tuple-ema} works for tuple-based ones. In this Thesis, we chose to work with tuple-based EMAs due to its simplicity, since it does not require keeping track of timestamp variations, but our approach can be adjusted to use time-based ones.

@@

In the next chapter, we review several outlier detection methods for time-series data and review state of the art sliding window aggregations and algorithms to summarize our data windows in a lighter representation.

### 2020-07-04 chapters/conclusions.tex
SAMM
Small corrections

In Section \ref{sec:hypothesis} we defined our hypothesis or main question to address as whether sliding window aggregations and outlier detection methods combined could be used to detect data pattern shifts in data streaming scenarios in the presence of high volumes of high velocity, highly skewed and seasonal data in real-time and with a low memory footprint. Before discussing our hypothesis, we revisit the research questions posed in Section \ref{sec:rqs} and provide answers to them.

@@

This research question really is about two things: maintaining the sliding window aggregation in real-time and keeping its memory usage constant.

The aggregation in question is an histogram aggregation.
An histogram requires storing a fixed-size list of bins and corresponding counts, regardless of the volume of data. This means that for ever growing volumes of data our memory consumption stays the same.

However, maintaining a sliding window histogram in real-time is an harder challenge. Traditionally, sliding window maintenance requires evicting the oldest element from the window when a new one is inserted and updating the aggregation state. We could maintain the histogram in constant time using a Subtract-On-Evict (SOE) algorithm (Algorithm \ref{pseudo:soe}): for each new event, we would evict the oldest one and decrement the bin it belonged to and insert the new event in the window while incrementing the bin it belongs to. However, using this approach, we violate our sublinear memory consumption, because now we have to keep the entire window contents in memory, hence we grow linearly regarding sliding window size. We instead propose an EMA-like approximate histogram aggregation, in Section \ref{sec:ema-hist}. An EMA-like histogram has constant memory complexity and can be maintained in real-time.

@@

\subsubsection*{(RQ3) How can we measure divergence between the reference and sliding window aggregations?}
Statistical tests that measure the similarity between two probability distributions are common in the literature. Tests like Kolmogorov–Smirnov, Wassertein or Jensen–Shannon (the one we used) provide a bounded measure of distance between two probability distributions, or histograms.

In brief, since the aggregation we use is an histogram, we can use statistical tests that measure similarity between probability distributions. We chose to use the Jensen–Shannon Divergence (JSD). The JSD takes as input two histograms and outputs a value between 0 and 1. The closer this value is to 0 the more similar the distributions are. On the other hand, a JSD value of 1 represents two different distributions. We chose to use the JSD metric due to its simplicity and leave as future work the usage of other statistical tests.

@@

In Section \ref{sec:sampling-batch}, we explain how we define the threshold and hence define when we raise an alert. In our batch phase, we make a number of samples and measure the JSD distance between the histogram aggregations of those samples and our reference histogram aggregation. We end up with a list of JSD values. Given a new JSD value, we check against the now known distribution of JSD values and alert if above a certain percentile. Note that defining a threshold percentile is easier, because we are defining a probability: alerting above the 99th percentile means alerting for hypothesis with 1\% probability.

@@

\subsubsection*{Hypothesis Discussion}
In our synthetic data experiments, we were able to detect all of the introduced anomalies. Additionally, we were able to do it achieving very high streaming throughputs. In the real data experiments, once again, we managed to achieve high troughputs but the produced alerts were not very accurate. We defer further investigation to future work because of the time constraints of this thesis but leave open the possiblity that the approximate aggregation value is more inaccurate than what we thought it would be, due to the fact that the EMA tail holds all past events.

@@

    \item an histogram aggregation based on Exponential Moving Averages, that replaces the entire sliding window contents by a lighter representation, resulting in a constant in time and space complexity sliding window aggregation

@@

First and foremost, we believe the next steps are to reason about the obtained results in the real data experiments and find the root cause for the existence of many and large alert periods. As possible hypothesis to test, we suggest testing with EMAs with higher smoothing factors, so that past events are forgotten faster; additionally, we suggest decreasing $\gamma$, which further increases the number of samples to make in the batch phase.

In this thesis we used the Jensen–Shannon Divergence (JSD) as statistical test to measure the distance or divergence between our reference and target sliding window aggregation. We chose to use the JSD metric due to its simplicity and because it was used in previous work successfully \cite{SAMM}. We believe that future work comprises testing new distances in the literature. In particular, we believe the Wassertein distance is worth trying next, and illustrate why in Figure \ref{fig:wasserteinvsjsd}.
\textcolor{red}{TODO: make this image better}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/wasserstein-vs-jsd.pdf}
      \caption{Wassertein vs Jensen–Shannon distances}
      \label{fig:wasserteinvsjsd}
    \end{center}
\end{figure}
The Wassertein distance measures the work required to transport the probability mass from one histogram to the other. This measure will be larger the further way the probability mass is. In Figure \ref{fig:wasserteinvsjsd} scenario, while in the JSD measure is the same for both cases, the Wassertein distance is larger in case 2 because the probability mass has to be transported further.

In this thesis we had to deal with the multiple test problem and apply a multiple test correction. In our case, we used Holm-Bonferroni due to its simplicity and because it controls the family wise error rate with a larger power than some other methods (thus providing a good trade-off between simplicity and power). There are many other multiple test correction methods, and we leave the testing of each as future work.

>>>>

In Section \ref{sec:hypothesis} we defined our hypothesis or main question to address as to whether sliding window aggregations and outlier detection methods combined could be used to detect data pattern shifts in data streaming scenarios in the presence of high volumes of high velocity, highly skewed and seasonal data in real-time and with a low memory footprint. Before discussing our hypothesis, we revisit the research questions posed in Section \ref{sec:rqs} and provide answers to them.

@@

This research question is about two things: maintaining the sliding window aggregation in real-time and keeping its memory usage constant.

The aggregation in question is a histogram aggregation.
A histogram requires storing a fixed-size list of bins and corresponding counts, regardless of the volume of data. This means that for ever-growing volumes of data our memory consumption stays the same.

However, maintaining a sliding window histogram in real-time is a harder challenge. Traditionally, sliding window maintenance requires evicting the oldest element from the window when a new one is inserted and updating the aggregation state. We could maintain the histogram in constant time using a Subtract-On-Evict (SOE) algorithm (Algorithm \ref{pseudo:soe}): for each new event, we would evict the oldest one and decrement the bin it belonged to and insert the new event in the window while incrementing the bin it belongs to. However, using this approach, we violate our sublinear memory consumption, because now we have to keep the entire window contents in memory, hence we grow linearly regarding sliding window size. We instead propose an EMA-like approximate histogram aggregation, in Section \ref{sec:ema-hist}. An EMA-like histogram has constant memory complexity and can be maintained in real-time.

@@

\subsubsection*{(RQ3) How can we measure divergence between the reference and sliding window aggregations?}
Statistical tests that measure the similarity between two probability distributions are common in the literature. Tests like Kolmogorov–Smirnov, Wasserstein or Jensen–Shannon (the one we used) provide a bounded measure of distance between two probability distributions, or histograms.

In brief, since the aggregation we use is a histogram, we can use statistical tests that measure the similarity between probability distributions. We chose to use the Jensen–Shannon Divergence (JSD). The JSD takes as input two histograms and outputs a value between 0 and 1. The closer this value is to 0 the more similar the distributions are. On the other hand, a JSD value of 1 represents two different distributions. We chose to use the JSD metric due to its simplicity and leave as future work the usage of other statistical tests.

@@

In Section \ref{sec:sampling-batch}, we explain how we define the threshold and hence define when we raise an alert. In our batch phase, we make a number of samples and measure the JSD distance between the histogram aggregations of those samples and our reference histogram aggregation. We end up with a list of JSD values. Given a new JSD value, we check against the now known distribution of JSD values and alert if above a certain percentile. Note that defining a threshold percentile is easier because we are defining a probability: alerting above the 99th percentile means alerting for hypotheses with 1\% probability.

@@

\subsubsection*{Hypothesis Discussion}
In our synthetic data experiments, we were able to detect all of the introduced anomalies. Additionally, we were able to do it achieving very high streaming throughputs. In the real data experiments, once again, we managed to achieve high throughput but the produced alerts were not very accurate. We defer further investigation to future work because of the time constraints of this thesis but leave open the possibility that the approximate aggregation value is more inaccurate than what we thought it would be since the EMA tail holds all past events.

@@

    \item a histogram aggregation based on Exponential Moving Averages, that replaces the entire sliding window contents by a lighter representation, resulting in a constant in time and space complexity sliding window aggregation

@@

First and foremost, we believe the next steps are to reason about the obtained results in the real data experiments and find the root cause for the existence of many and large alert periods. As a possible hypothesis to test, we suggest testing with EMAs with higher smoothing factors, so that past events are forgotten faster; additionally, we suggest decreasing $\gamma$, which further increases the number of samples to make in the batch phase.

In this thesis, we used the Jensen–Shannon Divergence (JSD) as a statistical test to measure the distance or divergence between our reference and target sliding window aggregation. We chose to use the JSD metric due to its simplicity and because it was used in previous work successfully \cite{SAMM}. We believe that future work comprises testing new distances in the literature. In particular, we believe the Wasserstein distance is worth trying next, and illustrate why in Figure \ref{fig:wasserteinvsjsd}.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/wasserstein-vs-jsd.pdf}
      \caption{Wassertein vs Jensen–Shannon distances}
      \label{fig:wasserteinvsjsd}
    \end{center}
\end{figure}
The Wasserstein distance measures the work required to transport the probability mass from one histogram to the other. This measure will be larger the further way the probability mass is. In Figure \ref{fig:wasserteinvsjsd} scenario, while in the JSD measure is the same for both cases, the Wasserstein distance is larger in case 2 because the probability mass has to be transported further.

In this thesis, we had to deal with the multiple test problem and apply a multiple test correction. In our case, we used Holm-Bonferroni due to its simplicity and because it controls the family-wise error rate with larger power than some other methods (thus providing a good trade-off between simplicity and power). There are many other multiple test correction methods, and we leave the testing of each as future work.

### 2020-07-04 chapters/introduction.tex
Small corrections

Thus the objective of this Thesis is to build a lightweight real-time system that monitors a stream of high volumes of high velocity, highly skewed and seasonal data and detects pattern shifts relative to a reference period.

\section{Motivation} \label{sec:motivation}
For any organization, the quality of their provided services is paramount. Therefore, it is common for companies to have auxiliary systems monitoring their services, testing their usage rate, availability and performance, just to name a few. The motivation behind this Thesis is to produce a system capable of real-time monitoring a stream of data so that changes in the underlying stream are caught early, preventing a fall in performance of the system analyzing said data stream.

@@

\section{Hypothesis} \label{sec:hypothesis}
The question this Thesis aims to answer is whether \textit{sliding window aggregations and outlier detection methods combined can be used to detect data pattern shifts in data streaming scenarios in the presence of high volumes of high velocity, highly skewed and seasonal data in real-time and with a low memory footprint.} In this Thesis, we study sliding window aggregations that replace the sliding window events in memory by a lighter representation, while still being able to provide an estimate of the aggregation value.

@@

In Chapter \ref{chap:sota}, we examine state of the art methods related with the previously mentioned topics and discuss shortcomings, strengths and the applicability of such methods to our hypothesis.

@@

In Chapter \ref{chap:validation}, we describe the datasets used and the set of experiments that were performed to validate our method. We start with single feature experiments on synthetic data, then multi feature analysis and finally move on to multi feature analysis on real data.

>>>>

Thus the objective of this thesis is to build a lightweight real-time system that monitors a stream of high volumes of high velocity, highly skewed and seasonal data and detects pattern shifts relative to a reference period.

\section{Motivation} \label{sec:motivation}
For any organization, the quality of their provided services is paramount. Therefore, it is common for companies to have auxiliary systems monitoring their services, testing their usage rate, availability and performance, just to name a few. The motivation behind this thesis is to produce a system capable of real-time monitoring a stream of data so that changes in the underlying stream are caught early, preventing a fall in the performance of the system analyzing said data stream.

@@

\section{Hypothesis} \label{sec:hypothesis}
The question this thesis aims to answer is whether \textit{sliding window aggregations and outlier detection methods combined can be used to detect data pattern shifts in data streaming scenarios in the presence of high volumes of high velocity, highly skewed and seasonal data in real-time and with a low memory footprint.} In this thesis, we study sliding window aggregations that replace the sliding window events in memory by a lighter representation, while still being able to provide an estimate of the aggregation value.

@@

In Chapter \ref{chap:sota}, we examine state of the art methods related to the previously mentioned topics and discuss shortcomings, strengths and the applicability of such methods to our hypothesis.

@@

In Chapter \ref{chap:validation}, we describe the datasets used and the set of experiments that were performed to validate our method. We start with single feature experiments on synthetic data, then multi-feature analysis and finally move on to multi-feature analysis on real data.

### 2020-07-04 chapters/method.tex
Small corrections

\subsubsection*{Multiple Test Correction} \label{sec:multi-test}
Note that we periodically apply this test (Equation \ref{eq:alert-test}) for each feature, effectively repeating the analysis. When considering several hypotheses, the problem of multiplicity arises: the more hypotheses are checked, the higher the probability of obtaining false positives. In other words, when repeating a test multiple times the multiple test or simultaneous statistical inference problem \cite{MultiTestProblem-Pubmed}
\cite{MultiTestProblem-Dickhaus2014}
\cite{MultiTestProblem-Miller1966}  \cite{multiple-test-correction-geoffrey} arises which states that the more inferences are made, the more likely erroneous inferences are to occur. In the literature, many multiple test correction methods are proposed, but we opted to use the Holm-Bonferroni multiple test correction because it is a simple and easy to compute correction that can reduce the amount of false negatives while having as target to control the amount of false positives. Furthermore, it does not require to assume any indepedence between the tested hypothesis (as other methods do) which would almost surely be a wrong assumption for realistic datasets. Therefore we leave as future work the testing of other correction methods. In statistics, Family-Wise Error Rate (FWER) \cite{MultivariateMT} \cite{MultitestTamhane2018AdvancesIP} is the probability of having one or more false positives when performing multiple hypotheses tests, \textit{i.e.}, considering one hypothesis true when it is not. The Holm-Bonferroni method \cite{HolmBonferroni} is one of many approaches for controlling the Family-Wise Error Rate \textit{(FWER)} by adjusting the rejection criteria for each hypothesis.

>>>>

\subsubsection*{Multiple Test Correction} \label{sec:multi-test}
Note that we periodically apply this test (Equation \ref{eq:alert-test}) for each feature, effectively repeating the analysis. When considering several hypotheses, the problem of multiplicity arises: the more hypotheses are checked, the higher the probability of obtaining false positives. In other words, when repeating a test multiple times the multiple test or simultaneous statistical inference problem \cite{MultiTestProblem-Pubmed}
\cite{MultiTestProblem-Dickhaus2014}
\cite{MultiTestProblem-Miller1966}  \cite{multiple-test-correction-geoffrey} arises which states that the more inferences are made, the more likely erroneous inferences are to occur. In the literature, many multiple test correction methods are proposed, but we opted to use the Holm-Bonferroni multiple test correction because it is a simple and easy to compute correction that can reduce the number of false negatives while having as target to control the number of false positives. Furthermore, it does not require to assume any independence between the tested hypothesis (as other methods do) which would almost surely be a wrong assumption for realistic datasets. Therefore we leave as future work the testing of other correction methods. In statistics, Family-Wise Error Rate (FWER) \cite{MultivariateMT} \cite{MultitestTamhane2018AdvancesIP} is the probability of having one or more false positives when performing multiple hypotheses tests, \textit{i.e.}, considering one hypothesis true when it is not. The Holm-Bonferroni method \cite{HolmBonferroni} is one of many approaches for controlling the Family-Wise Error Rate \textit{(FWER)} by adjusting the rejection criteria for each hypothesis.
